---
title: "Regrid Checks and Column Integrated Plots"
author: "Victoria Froh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Introduction

1.  Checking regrid subregion
2.  Verifying scope of dTA in regrid
3.  Column-integrated dTA maps

```{r setup, message=FALSE}

#loading packages
library(tidyverse)
library(data.table)
library(arrow)
library(scales)
library(maps)
library(geosphere)

# Path to intermediate computation outputs
path_outputs <- "/net/sea/work/vifroh/oae_ccs_roms_data/regrid_2/"

# Path to save practice plots when working on them
path_plots <- "/net/sea/work/vifroh/test_plots/"

# loading in dTA conc data to make a surface plot/columnint
lanina_dTA_conc <- read_feather(
  paste0(path_outputs,"lanina_dTA_concdataRG2.feather"))
neutral_dTA_conc <- read_feather(
  paste0(path_outputs,"neutral_dTA_concdataRG2.feather"))
elnino_dTA_conc <- read_feather(
  paste0(path_outputs,"elnino_dTA_concdataRG2.feather"))

# loading in dTA full integration data for competency check
lanina_dTA_int<- read_feather(
  paste0(path_outputs,"lanina_CDReff_intRG2.feather"))

# loading in dTA sum original grid data for competency check
lanina_intdata_ogs <- read_feather(
  "/net/sea/work/vifroh/oae_ccs_roms_data/regrid/lanina_dTAint_comparegrids.feather")

```

# Looking at the boundaries of the regridded subregion
```{r subset_plot, message=FALSE, warning=FALSE}

# filtering dTA conc data to make a surface plot
surface_data <- lanina_dTA_conc[depth == 0 & time == "1998-09"]

# Convert lat and lon to numeric 
surface_data$lat <- as.numeric(surface_data$lat)
surface_data$lon <- as.numeric(surface_data$lon)

# # Convert longitude to -180 to 180 range 
# surface_data$lon <- surface_data$lon - 360

# Define the bounding box for the plot
lat_range <- range(surface_data$lat, na.rm = TRUE)
lon_range <- range(surface_data$lon, na.rm = TRUE)

# plotting surface map
ggplot() + 
  geom_polygon(data = map_data("world"), aes(x = long, y = lat, group = group),
               fill = "lightgray", color = "white") +
  geom_raster(data = surface_data, aes(x = lon, y = lat, fill = dTA)) +
  scale_fill_viridis_c() +  # Change the color scale to suit your data
  theme_minimal() +
  coord_fixed(xlim = c(lon_range[1] - 2, lon_range[2] + 2),
              ylim = c(lat_range[1] - 2, lat_range[2] + 2)) + 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8),
    axis.text.y = element_text(size = 8),
    axis.ticks = element_line(size = 0.5),
    panel.grid = element_blank()
  ) +
  labs(title = "Subset Domain (La Niña, September 1998)",
       x = "Longitude",
       y = "Latitude",
       fill = "dTA (mmol/m^3)")

# # save plot
# ggsave(paste0(path_plots, "regrid_domainRG2.png"), plot = last_plot(), 
#        width = 8, height = 6, dpi = 300)

rm(surface_data)
gc()

```

## Checking the alkalinity containment within the regridded subregion over the time series
```{r scope_check, message=FALSE, eval=FALSE}

# using full integrated data from cdr_eff_molar file
setDT(lanina_dTA_int)

# combine with original dTA sum data to compare; subtracting regrid from original
lanina_dTAint_compare <- merge(lanina_intdata_ogs, lanina_dTA_int[, .(time, dTA_sum_rg2 = dTA_sum)],
                          by = "time", all.x = FALSE) %>% 
  .[, dTA_dif_rg2 := dTA_sum_og - dTA_sum_rg2] %>% 
  .[, frac_miss_rg2 := dTA_dif_rg2 / dTA_sum_og]

# # save data file
# write_feather(lanina_dTAint_compare, paste0(path_outputs,
#                                       "lanina_dTAint_comparegridsRG2.feather"))

rm(lanina_dTAint_compare, lanina_intdata_ogs)
gc()

```

# Column integrated plots of added alkalinity 
To check lateral movement during the time series

```{r subregions, eval=FALSE}
# Defining Subregions (currently abandoned for now)

# loc_box <- c(30, 35, -115, -122)
# nep_box <- c(10, 60, -155, -95)
#  #work in progress
# coastline <- map("world", plot = FALSE, fill = FALSE,
#                  xlim = c(-130, -113), ylim = c(20, 50))
# coastline <- data.table(lon = coastline$x, lat = coastline$y)[!is.na(lon) & !is.na(lat)]
# coastline <- coastline[lon >= -130 & lon <= -113 & lat >= 20 & lat <= 50]
# 
# # filtering
# 
# 
# rm()
# gc()

```

## Column Integrated Plots
```{r column_integrated_plots, message=FALSE}
# rerunning these on own for different phases/months; can load saved column int
# files from folder

# using conc data then multiplied by depth bin size so have mmol/m^2
setDT(lanina_dTA_conc)
lanina_dTA_conc$depth <- as.numeric(lanina_dTA_conc$depth)
# filter out only top 100m for CDReff integrated plot
lanina_dTA_conc <- lanina_dTA_conc[depth <= 100]

lanina_dTA_conc <-
  lanina_dTA_conc[, thickness :=
                    ifelse(depth == 0, 2.5,
                           ifelse(depth < 80, 5,
                                  ifelse(depth == 80, 7.5,
                                         ifelse(depth < 100, 10, #edit here for top 100
                                               5
                                                # ifelse(depth == 100, 15,
                                                #        ifelse(depth < 300, 20,
                                                #               10
                                                ))))
                  ] # two )) removed for top100
lanina_dTA_conc <- lanina_dTA_conc[, dTA_m2 := dTA * thickness] %>%  # units now moles/m2
      .[, dDIC_m2 := dDIC * thickness] %>% 
      .[, CDReff_m2 := fifelse(dDIC_m2/dTA_m2 == Inf, NaN, dDIC_m2/dTA_m2)] 
# calculating CDR Efficiency per cell, replacing dDIC calcs producing Inf -> Na

# grouping by lat/lon and integrating vertically, averaging the CDR Efficiency
lanina_dTA_columnint<- lanina_dTA_conc[, .(dTA_column = sum(dTA_m2, na.rm = TRUE),
                                           dDIC_column = sum(dDIC_m2, na.rm = TRUE),
                                           CDReff_avg = mean(CDReff_m2, na.rm = TRUE)), 
                                       # averaging CDReff of individual grid cells
                                      by = c("lat", "lon", "time")] %>% 
 .[, CDReff_col := fifelse(dDIC_column/dTA_column == Inf, NaN, dDIC_column/dTA_column)] 
# total CDReff for column; this is fine for top 100m if thinking this is mixing

# Convert lat, lon to numeric 
lanina_dTA_columnint$lat <- as.numeric(lanina_dTA_columnint$lat)
lanina_dTA_columnint$lon <- as.numeric(lanina_dTA_columnint$lon)

# filter by time to create a timestop plot
surface_data <- lanina_dTA_columnint[time == "2000-05"]

# # Convert longitude to -180 to 180 range 
# surface_data$lon <- surface_data$lon - 360

# plotting column integrated map
ggplot() + 
  geom_polygon(data = map_data("world"), aes(x = long, y = lat, group = group),
               fill = "lightgray", color = "white") +
  geom_raster(data = surface_data, aes(x = lon, y = lat, fill = CDReff_col)) +
  scale_fill_viridis_c(limit = c(0, 1)) +  # set the color range
  theme_light() +
  coord_fixed(xlim = c(-170, -85),
              ylim = c(10, 60)) +
  scale_x_continuous(breaks = seq(-170, -85, by = 10)) +
  scale_y_continuous(breaks = seq(10, 60, by = 10)) +
  labs(title = "Vertically Integrated CDR Efficiency, top 100m (La Niña, May 2000)", # La Niña
       x = "Longitude",
       y = "Latitude",
       fill = "CDR Efficiency") + # dTA (mmol/m^2)
  theme(panel.border = element_blank())

# # save plot
# ggsave(paste0(path_plots, "lanina_columnint_May2000_CDReff100.png"), plot = last_plot(),
#        width = 8, height = 6, dpi = 300)

# save column integrated data
# write_feather(lanina_dTA_columnint, paste0(path_outputs,
#                                       "lanina_columnintRG2_top100.feather"))
# write_feather(neutral_dTA_columnint, paste0(path_outputs,
#                                        "neutral_columnintRG2_top100.feather"))
# write_feather(elnino_dTA_columnint, paste0(path_outputs,
#                                       "elnino_columnintRG2_top100.feather"))

rm(lanina_dTA_columnint, surface_data, lanina_dTA_conc)
gc()

```

## Difference From Mean Plots
```{r column_intdfm_plots, message=FALSE}

# load in full table with column ints and mean
colint_data <- read_feather(paste0(path_outputs, "colint_RG2.feather"))

# calc dfm

# dTA
col_dt <- colint_data[colint_data$phase == "neutral" & colint_data$month == 10]
ggplot() +
  geom_polygon(data = map_data("world"), aes(x = long, y = lat, group = group),
               fill = "lightgray", color = "white") +
  geom_raster(data = col_dt, aes(x = lon, y = lat, fill = dTA_column - dTA_mean)) +
  geom_rect(aes(xmin = -118.0625, xmax = -117.9375,
                ymin = 33.5625, ymax = 33.6875),
            fill = NA, color = "black", size = 0.5) +  # Outline only (no fill)
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,
                         guide = guide_colorbar(
      barwidth = unit(0.7, "cm"), barheight = unit(6, "cm"), title.hjust = 0.5)#,
      #limits = c(-200000000, 200000000) 
)+
  # scale_fill_viridis_c(limit = c(-0.5, 103), guide = guide_colourbar(
  #   barwidth = 15, barheight = 0.5, title.position = "bottom", title.hjust = 0.5
  # )) +  # set the color range
  theme_light() +
  coord_fixed(xlim = c(-125, -115), #nep grid
              ylim = c(25, 35)) +
  # coord_fixed(xlim = c(-127, -115), # loc grid
  #             ylim = c(27.5, 37.5)) +
  scale_x_continuous(breaks = seq(-155, -85, by = 5)) +
  scale_y_continuous(breaks = seq(10, 60, by = 5)) +
  labs(#title = "Added Alkalinity Mean Column Concentration (Month 13)",
       x = "Longitude",
       y = "Latitude",
       fill = "\u0394TA' [mmol/m\u00B2]") + # dTA (mmol/m^2)
  theme(panel.border = element_blank(),
        axis.text = element_text(size = 11, color = "black"),
        axis.title = element_text(size = 12),
        axis.ticks = element_line(color = "black", size = 0.5),
        axis.ticks.length = unit(0.2, "cm"),
        legend.text = element_text(size = 9),
        legend.title = element_text(size = 10)#,
        #legend.position = "bottom",
        #legend.direction = "horizontal"
        )

# # save plot
# ggsave(paste0(path_plots, "dTA_col_mean_10mo_neutdfm.png"), plot = last_plot(),
#        width = 6, height = 6, dpi = 300)

rm(list = ls())
gc()

```
