---
title: "Drivers - Mixing Depth"
author: "Victoria Froh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

1. Calculate and map maximum mixing depth of each grid cell over first 2 years (all 3 phases)
2. Map the difference from the 3-phase mean MMD for each grid cell
3. Map the regional distribution of mixing depth in select months (all 3 phases)
4. Map the difference from the 3-phase mean MD per grid cell for select months
5. Load in dTA data and calculate what fraction was beneath the MMD for time series
6. Repeat subducted time series for dDIC data

```{r setup, message=FALSE}

#loading packages
library(tidyverse)
library(tidync)
library(data.table)
library(arrow)
library(scales)

# Path to intermediate computation outputs
path_outputs <- "/net/sea/work/vifroh/oae_ccs_roms_data/regrid_2/"

# Path to save practice plots when working on them
path_plots <- "/net/sea/work/vifroh/test_plots/"

# loading in previous saved surface variable data
lanina_surf_data <- read_feather(
  paste0(path_outputs, "lanina_surf_dataRG2.feather"))
neutral_surf_data <- read_feather(
  paste0(path_outputs, "neutral_surf_dataRG2.feather"))
elnino_surf_data <- read_feather(
  paste0(path_outputs, "elnino_surf_dataRG2.feather"))

# loading in dTA molar data
lanina_dTA_mol <- read_feather(
  paste0(path_outputs,"lanina_dTA_moldataRG2.feather"))
neutral_dTA_mol <- read_feather(
  paste0(path_outputs,"neutral_dTA_moldataRG2.feather"))
elnino_dTA_mol <- read_feather(
  paste0(path_outputs,"elnino_dTA_moldataRG2.feather"))

```

```{r data_prep, message=FALSE}

phase_data <- list(lanina_surf_data, neutral_surf_data, elnino_surf_data)

# Convert lat/lon to numeric and subset first two years
phase_data <- phase_data %>%
  lapply(function(table) {
    setDT(table)
    setorder(table, time)
    table[, lat := as.numeric(lat)]
    table[, lon := as.numeric(lon)]
    table[, month := .GRP, by = time] # gives index in order to each unique time
    table <- table[month <= 24] # subsets to first two years
    
    return(table)
  })

# return tables to each item 
lanina_surf_data <- phase_data[[1]]
neutral_surf_data <- phase_data[[2]]
elnino_surf_data <- phase_data[[3]]

# for alkalinity subduction: using full grid dTA moles data, subset first 2 years
mol_data <- list(lanina_dTA_mol, neutral_dTA_mol, elnino_dTA_mol)

mol_data <- mol_data %>%
  lapply(function(table) {
    setDT(table)
    setorder(table, time)
    table[, month := .GRP, by = time] # gives index in order to each unique time
    table <- table[month <= 24] # subsets to first two years
    
    return(table)
  })

# return tables to each item 
lanina_dTA_mol <- mol_data[[1]]
neutral_dTA_mol <- mol_data[[2]]
elnino_dTA_mol <- mol_data[[3]]

# plotting names for later
phase_titles <- c("La Niña", "Neutral", "El Niño")
phases <- c("lanina", "neutral", "elnino")

```

## Exploring Mixing Depth

Regional distribution of maximum depths

```{r hbls_max, message=FALSE}

# finding max mixing depth at each grid cell
phase_data <- phase_data %>%
  lapply(function(table) {
    table[, c("FG_CO2_OAE", "FG_CO2", "dFG") := NULL]
    # removes fugacity columns
    max_table <- table[, .SD[which.max(hbls)], by = .(lat, lon)] 
    # subsets by each lat/lon combo, calculates max hbls, keeps that row
    
    return(max_table)
  })

# return tables to new item
lanina_hblsmax <- phase_data[[1]]
neutral_hblsmax <- phase_data[[2]]
elnino_hblsmax <- phase_data[[3]]

# adding phase and combining all 3 into one table
phase_list <- Map(function(table, phase) {
    table[, phase := phase]
  }, phase_data, phases)

hblsmax_data <- rbindlist(phase_list)

# plotting map of deepest hbls for each point
create_hblsmax_map <- function(phase_name, title_text) {
  phase_dt <- hblsmax_data[hblsmax_data$phase == phase_name,]
  plot <- ggplot(phase_dt, aes(x = lon, y = lat, fill = hbls)) +
    geom_polygon(data = map_data("world"), aes(x = long, y = lat, group = group), 
                 fill = "lightgray", color = "white") +
    geom_raster() +
    scale_fill_viridis_c(limit = c(0, 180)) +  # set the color range
    theme_light() +
    coord_fixed(xlim = c(-170, -85),
              ylim = c(10, 60)) +
    scale_x_continuous(breaks = seq(-170, -85, by = 10)) +
    scale_y_continuous(breaks = seq(10, 60, by = 10)) +
    labs(x = "Longitude",
         y = "Latitude",
         fill = "Maximum Mixing\nDepth (m)",
         title = paste0(title_text, " Phase, Maximum Mixing Depth in First Two Years")) +
    theme(panel.border = element_blank())
  
  print(plot)
  
  # # save plot
  # ggsave(paste0(path_plots, phase_name, "_hblsmax_map.png"), plot = plot,
  #        width = 8, height = 6, dpi = 300)
}

lapply(seq_along(phases), function(i) {
  create_hblsmax_map(phases[i], phase_titles[i])
})

rm(create_hblsmax_map)
gc()

```

Maximum depth difference from mean

```{r hbls_max_means, message=FALSE}

# calculating mean max mixing depth across all 3 phases for each grid cell
hblsmax_data[, ":=" (
  hbls_mean = mean(hbls, na.rm = TRUE)
), by = .(lat, lon)]

# plotting map of deepest hbls for each point
create_hblsmax_meanmap <- function(phase_name, title_text) {
  phase_dt <- hblsmax_data[hblsmax_data$phase == phase_name,]
  plot <- ggplot(phase_dt, aes(x = lon, y = lat, fill = hbls - hbls_mean)) +
    geom_polygon(data = map_data("world"), aes(x = long, y = lat, group = group), 
                 fill = "lightgray", color = "white") +
    geom_raster() +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,
                         limits = c(-50, 52)) +
    theme_light() +
    coord_fixed(xlim = c(-170, -85),
              ylim = c(10, 60)) +
    scale_x_continuous(breaks = seq(-170, -85, by = 10)) +
    scale_y_continuous(breaks = seq(10, 60, by = 10)) +
    labs(x = "Longitude",
         y = "Latitude",
         fill = "Maximum Mixing\nDepth Difference (m)",
         title = paste0(title_text, " Phase, Maximum Mixing Depth in First Two Years Difference from the Mean")) +
    theme(panel.border = element_blank())
  
  print(plot)
  
  # # save plot
  # ggsave(paste0(path_plots, phase_name, "_hblsmax_meanmap.png"), plot = plot,
  #        width = 8, height = 6, dpi = 300)
}

lapply(seq_along(phases), function(i) {
  create_hblsmax_meanmap(phases[i], phase_titles[i])
})

rm(create_hblsmax_meanmap)
gc()

```

Mixing depth maps during months of interest

```{r hbls_dist, message=FALSE}

phase_data <- list(lanina_surf_data, neutral_surf_data, elnino_surf_data)

# adding phase and combining all 3 into one table
phase_list <- Map(function(table, phase) {
    table[, phase := phase]
  }, phase_data, phases)

hblsfull_data <- rbindlist(phase_list) # now has all hbls data for all phases

# can rerun code for months of choice; 12-16
create_hbls_distmap <- function(phase_name, title_text) {
  phase_dt <- hblsfull_data[hblsfull_data$phase == phase_name & hblsfull_data$month == 22]
  plot <- ggplot(phase_dt, aes(x = lon, y = lat, fill = hbls)) +
    geom_polygon(data = map_data("world"), aes(x = long, y = lat, group = group), 
                 fill = "lightgray", color = "white") +
    geom_raster() +
    scale_fill_viridis_c(limit = c(0, 180)) +  # set the color range
    theme_light() +
    coord_fixed(xlim = c(-170, -85),
              ylim = c(10, 60)) +
    scale_x_continuous(breaks = seq(-170, -85, by = 10)) +
    scale_y_continuous(breaks = seq(10, 60, by = 10)) +
    labs(x = "Longitude",
         y = "Latitude",
         fill = "Mixing Depth (m)",
         title = paste0(title_text, " Phase, Mixing Depth (Month 22)")) +
    theme(panel.border = element_blank())
  
  print(plot)

  # # save plot
  # ggsave(paste0(path_plots, phase_name, "_hblsmap_month22.png"), plot = plot,
  #        width = 8, height = 6, dpi = 300)
}

lapply(seq_along(phases), function(i) {
  create_hbls_distmap(phases[i], phase_titles[i])
})

# write_feather(hblsfull_data, paste0(path_outputs,
#                                       "hblsfull_data.feather"))

rm(create_hbls_distmap)
gc()

```

Mixing depth difference from mean maps during months of interest

```{r hbls_dist_mean, message=FALSE}

# calculating mean mixing depth across all 3 phases for each grid cell, 1st 2 years
setDT(hblsfull_data)
hblsfull_data[, ":=" (
  hbls_mean = mean(hbls, na.rm = TRUE)
), by = .(lat, lon, month)]

# can rerun code for months of choice; 12-16
create_hbls_meanmap <- function(phase_name, title_text) {
  phase_dt <- hblsfull_data[hblsfull_data$phase == phase_name & hblsfull_data$month == 10]
                             # & lat > 20 & lat < 40 & lon > -130 & lon < -110]
  plot <- ggplot(phase_dt, aes(x = lon, y = lat, fill = hbls - hbls_mean)) +
    geom_polygon(data = map_data("world"), aes(x = long, y = lat, group = group), 
                 fill = "lightgray", color = "white") +
    geom_raster() +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,
                         limits = c(-75, 75), breaks = c(-50, -25, 0, 25, 50)) +
    theme_light() +
    coord_fixed(xlim = c(-170, -85),
              ylim = c(10, 60)) +
    scale_x_continuous(breaks = seq(-170, -85, by = 10)) +
    scale_y_continuous(breaks = seq(10, 60, by = 10)) +
    labs(x = "Longitude",
         y = "Latitude",
         fill = "Difference from Mean Mixing Depth (m)",
         title = paste0(title_text, " Phase, Mixing Depth Difference from Mean (Month 10)")) +
    theme(panel.border = element_blank())
  
  print(plot)
  
  # # save plot
  # ggsave(paste0(path_plots, phase_name, "_hblsmapdfm_month10.png"), plot = plot,
  #        width = 8, height = 6, dpi = 300)
}

lapply(seq_along(phases), function(i) {
  create_hbls_meanmap(phases[i], phase_titles[i])
})

# write_feather(hblsfull_data, paste0(path_outputs,
#                                       "hblsfull_meandata.feather"))

rm(create_hbls_meanmap)
gc()

```

Fraction of dTA subducted under MMD

```{r dta_sub_hblsmax, message=FALSE}

# merging relevant alk and hbls data
mol_data <- Map(function(mol_table, phase) {
  mol_table <- mol_table[, .SD, .SDcols = c("lon", "lat", "depth", "time", "month", 
                                      "dTA_full", "dDIC_full")]
  hbls_table <- get(paste0(phase, "_hblsmax")) # fetch matching hbls table
  new_table <- merge(mol_table, hbls_table[, .(lat, lon, hbls)],
                       by = c("lon", "lat"), all.x = TRUE) # merging
  return(new_table)
  }, mol_data, phases)

# return tables to each item, now mol data has the max mixing depth for each cell
lanina_dTA_mol <- mol_data[[1]]
neutral_dTA_mol <- mol_data[[2]]
elnino_dTA_mol <- mol_data[[3]]

# # setting boundaries of depth bins; only need if going to use boundaries to separate

# depth_ranges <- unique(lanina_dTA_mol[, .(depth)])
# setorder(depth_ranges, depth)
# 
# depth_ranges[, ":=" (
#     depth_upper = (shift(depth, type = "lag", fill = 0) + depth) / 2, # set shallower boundary
#     depth_lower = (shift(depth, type = "lead", fill = max(depth)) + depth) / 2 
#   )] # deeper bound
# 
# # merge depth ranges back to tables
# mol_data <- mol_data %>%
#   lapply(function(table) {
#     table <- merge(table, depth_ranges, by = "depth", all.x = TRUE)
#     return(table)
#   })

# # return tables to each item 
# lanina_dTA_mol <- mol_data[[1]]
# neutral_dTA_mol <- mol_data[[2]]
# elnino_dTA_mol <- mol_data[[3]]

# calculating subducted portion; assuming it stays subducted? would it make sense
# here to use the mixing depth at each month rather than the max?
subducted_data <- lapply(mol_data, function(table){
  new_table <- table[, .(dTA_int = sum(dTA_full, na.rm = TRUE), # full int dTA moles/month
            dTA_sub_int = sum(dTA_full[depth > hbls], na.rm = TRUE)), 
        # only the dTA subducted below MMD, approx with depth ie middle of bin
            by = .(month)]
  new_table[, frac_sub := dTA_sub_int / dTA_int] 
  # calculates approx portion subducted under each cell's 2-year max hbls 
  return(new_table)
  })

# return tables to each item 
lanina_dTA_subhbls <- subducted_data[[1]]
neutral_dTA_subhbls <- subducted_data[[2]]
elnino_dTA_subhbls <- subducted_data[[3]]

# combining all three tables in to one for ease of plotting
sub_list <- Map(function(table, phase) {
    table[, phase := phase]
  }, subducted_data, phases)

sub_fulldata <- rbindlist(sub_list)

# Plotting time series of all three
ggplot(sub_fulldata, aes(x = month, y = frac_sub, color = phase)) +
  scale_x_continuous(breaks = seq(0, 24, by = 3), limits = c(0,24),
                     expand = c(0,0)) +
  scale_y_continuous(limits = c(0, 0.4), expand = c(0,0)) +
  geom_vline(xintercept = seq(0, 24, by = 12), color = "grey75", linewidth = 0.4) +
  geom_line(linewidth = 1) +
  labs(title = "Added Alkalinity Subducted Beneath Maximum Mixing Depth", 
       x = "Months Since OAE Start", 
       y = "Fraction of Alkalinity Subducted") +
  theme_bw()

# ggsave(paste0(path_plots, "dTA_sub_hblsmax.png"), plot = last_plot(),
#               width = 10, height = 6, dpi = 300)

rm(depth_ranges, sub_list, subducted_data)
gc()
```


```{r ddic_sub_hblsmax, message=FALSE}

# calculating subducted portion; assuming it stays subducted? would it make sense
# here to use the mixing depth at each month rather than the max?
subducted_data <- lapply(mol_data, function(table){
  new_table <- table[, .(dDIC_int = sum(dDIC_full, na.rm = TRUE), # full int dDIC moles/month
            dDIC_sub_int = sum(dDIC_full[depth > hbls], na.rm = TRUE)), 
        # only the dTA subducted below MMD, approx with depth ie middle of bin
            by = .(month)]
  new_table[, frac_sub := dDIC_sub_int / dDIC_int] 
  # calculates approx portion subducted under each cell's 2-year max hbls 
  return(new_table)
  })

# return tables to each item 
lanina_dDIC_subhbls <- subducted_data[[1]]
neutral_dDIC_subhbls <- subducted_data[[2]]
elnino_dDIC_subhbls <- subducted_data[[3]]

# combining all three tables in to one for ease of plotting
sub_list <- Map(function(table, phase) {
    table[, phase := phase]
  }, subducted_data, phases)

sub_fulldata <- rbindlist(sub_list)

# Plotting time series of all three, either fraction or moles
ggplot(sub_fulldata, aes(x = month, y = frac_sub, color = phase)) +
  scale_x_continuous(breaks = seq(0, 24, by = 3), limits = c(0,24),
                     expand = c(0,0)) +
  scale_y_continuous(limits = c(0, 0.4), expand = c(0,0)) +
  geom_vline(xintercept = seq(0, 24, by = 12), color = "grey75", linewidth = 0.4) +
  geom_line(linewidth = 1) +
  labs(title = "New DIC Subducted Beneath Maximum Mixing Depth", 
       x = "Months Since OAE Start", 
       y = "Fraction of dDIC Subducted") +
  theme_bw()

# ggsave(paste0(path_plots, "dDIC_sub_hblsmax.png"), plot = last_plot(),
#               width = 10, height = 6, dpi = 300)

rm(list = ls())
gc()
```

