---
title: "Drivers - Air-Sea CO2 Flux"
author: "Victoria Froh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

1. Map air-sea CO2 flux of each surface cell (all 3 phases, monthly maps)
2. Map dFlux of each cell (OAE - control), monthly maps, also on zoomed in region
3. Map dFlux difference from mean of each phase on zoomed region
4. Calculate moles of CO2 uptake per month dist and time series


```{r setup, message=FALSE}

#loading packages
library(tidyverse)
library(tidync)
library(data.table)
library(arrow)
library(scales)

# Path to intermediate computation outputs
path_outputs <- "/net/sea/work/vifroh/oae_ccs_roms_data/regrid_2/"

# Path to original grid intermediate computation outputs
path_outputs_og <- "/net/sea/work/vifroh/oae_ccs_roms_data/original_grid/"

# Path to save practice plots when working on them
path_plots <- "/net/sea/work/vifroh/test_plots/"

# loading in previous saved surface variable data
lanina_surf_data <- read_feather(
  paste0(path_outputs, "lanina_surf_dataRG2.feather"))
neutral_surf_data <- read_feather(
  paste0(path_outputs, "neutral_surf_dataRG2.feather"))
elnino_surf_data <- read_feather(
  paste0(path_outputs, "elnino_surf_dataRG2.feather"))

# loading in previous saved data with area
lanina_dTA_mol <- read_feather(
  paste0(path_outputs,"lanina_dTA_mol_mdataRG2.feather"))

# loading in original grid integrated CDR data
lanina_cdreff_sum <- read_feather(
  paste0(path_outputs_og, "lanina_CDReff_integrated.feather"))
neutral_cdreff_sum <- read_feather(
  paste0(path_outputs_og, "neutral_CDReff_integrated.feather"))
elnino_cdreff_sum <- read_feather(
  paste0(path_outputs_og, "elnino_CDReff_integrated.feather"))

```

```{r data_prep, message=FALSE}

phase_data <- list(lanina_surf_data, neutral_surf_data, elnino_surf_data)

# Convert lat/lon to numeric and subset first two years
phase_data <- phase_data %>%
  lapply(function(table) {
    setDT(table)
    setorder(table, time)
    table[, lat := as.numeric(lat)]
    table[, lon := as.numeric(lon)]
    table[, month := .GRP, by = time] # gives index in order to each unique time
    table <- table[month <= 24] # subsets to first two years
    
    return(table)
  })

# return tables to each item 
lanina_surf_data <- phase_data[[1]]
neutral_surf_data <- phase_data[[2]]
elnino_surf_data <- phase_data[[3]]

# plotting names for later
phase_titles <- c("La Niña", "Neutral", "El Niño")
phases <- c("lanina", "neutral", "elnino")

# adding phase and combining all 3 into one table for plotting
phase_list <- Map(function(table, phase) {
    table[, phase := phase]
  }, phase_data, phases)

surface_data <- rbindlist(phase_list)

# subset down to just area data; figure out how to subset just the unique?
setDT(lanina_dTA_mol)
lanina_dTA_mol <- unique(lanina_dTA_mol[, .(lat, lon, area)])

# create combined table of original grid data with "phase" column
full_cdreff_og <- rbindlist(
  list(lanina = lanina_cdreff_sum, neutral = neutral_cdreff_sum,
       elnino = elnino_cdreff_sum),
  idcol = "phase")

# add column for months since oae addition begins for x-axis
full_cdreff_og <- full_cdreff_og[, month := seq_len(.N), by = phase]

rm(elnino_cdreff_sum, lanina_cdreff_sum, neutral_cdreff_sum)
gc()

```

## Exploring Air-Sea FLux

Regional distribution of flux in OAE runs

```{r fgCO2_dist, message=FALSE}

# can rerun code for months of choice
create_fgOAE_distmap <- function(phase_name, title_text) {
  phase_dt <- surface_data[surface_data$phase == phase_name & surface_data$month == 14]
  plot <- ggplot(phase_dt, aes(x = lon, y = lat, fill = FG_CO2_OAE)) +
    geom_polygon(data = map_data("world"), aes(x = long, y = lat, group = group), 
                 fill = "lightgray", color = "white") +
    geom_raster() +
    scale_fill_distiller(limit = c(-0.0055, 0.0055), palette = "PRGn") +  # set the color range
    theme_light() +
    coord_fixed(xlim = c(-170, -85),
              ylim = c(10, 60)) +
    scale_x_continuous(breaks = seq(-170, -85, by = 10)) +
    scale_y_continuous(breaks = seq(10, 60, by = 10)) +
    labs(x = "Longitude",
         y = "Latitude",
         fill = "Air-Sea CO2 Flux (mmol/m^2/s)",
         title = paste0(title_text, " Phase, Air-Sea CO2 Flux (Month 14)")) +
    theme(panel.border = element_blank())
  
  print(plot)

  # # save plot
  # ggsave(paste0(path_plots, phase_name, "_fgOAEmap_month14.png"), plot = plot,
  #        width = 8, height = 6, dpi = 300)
}

lapply(seq_along(phases), function(i) {
  create_fgOAE_distmap(phases[i], phase_titles[i])
})

rm(create_fgOAE_distmap, phase_data, phase_list)
gc()

```

Regional distribution of dflux in OAE runs (OAE flux - control flux)

```{r dfgCO2_dist, message=FALSE}

# can rerun code for months of choice,
create_dfg_distmap <- function(phase_name, title_text) {
  phase_dt <- surface_data[surface_data$phase == phase_name & surface_data$month == 22]
  plot <- ggplot(phase_dt, aes(x = lon, y = lat, fill = dFG)) +
    geom_polygon(data = map_data("world"), aes(x = long, y = lat, group = group), 
                 fill = "lightgray", color = "white") +
    geom_raster() +
    # scale_fill_viridis_c(limit = c(0, 0.00022)) +
    scale_fill_gradient2(low = "green", mid = "white", high = "purple", midpoint = 0,
                         limit = c(0, 0.00022)) +
    theme_light() +
    coord_fixed(xlim = c(-170, -85),
              ylim = c(10, 60)) +
    scale_x_continuous(breaks = seq(-170, -90, by = 10)) +
    scale_y_continuous(breaks = seq(10, 60, by = 10)) +
    labs(x = "Longitude",
         y = "Latitude",
         fill = "dCO2 Flux 
         (mmol/m^2/s)",
         title = paste0(title_text, " Phase, Change in Air-Sea CO2 Flux (Month 22)")) +
    theme(panel.border = element_blank())
  
  print(plot)

  # # save plot
  # ggsave(paste0(path_plots, phase_name, "_dfgmap_month22.png"), plot = plot,
  #        width = 8, height = 6, dpi = 300)
}

lapply(seq_along(phases), function(i) {
  create_dfg_distmap(phases[i], phase_titles[i])
})

rm(create_dfg_distmap)
gc()

```

Zoomed in maps of dflux in OAE runs (OAE flux - control flux) near add site

```{r dfgCO2_dist_zoom, message=FALSE}

# can rerun code for months of choice
create_dfg_distmap <- function(phase_name, title_text) {
  phase_dt <- surface_data[surface_data$phase == phase_name & surface_data$month == 15]
  plot <- ggplot(phase_dt, aes(x = lon, y = lat, fill = dFG)) +
    geom_polygon(data = map_data("world"), aes(x = long, y = lat, group = group), 
                 fill = "lightgray", color = "white") +
    geom_raster() +
    # scale_fill_viridis_c(limit = c(0, 0.00022)) +
    scale_fill_gradient2(low = "green", mid = "white", high = "purple", midpoint = 0,
                         limit = c(0, 0.0000575)) +
    theme_light() +
    coord_fixed(xlim = c(-135, -100),
              ylim = c(10, 45)) +
    scale_x_continuous(breaks = seq(-135, -100, by = 10)) +
    scale_y_continuous(breaks = seq(10, 45, by = 10)) +
    labs(x = "Longitude",
         y = "Latitude",
         fill = "dCO2 Flux 
         (mmol/m^2/s)",
         title = paste0(title_text, " Phase, Change in Air-Sea CO2 Flux (Month 15)")) +
    theme(panel.border = element_blank())
  
  print(plot)

  # # save plot
  # ggsave(paste0(path_plots, phase_name, "_dfgmapzoom_month15.png"), plot = plot,
  #        width = 8, height = 6, dpi = 300)
}

lapply(seq_along(phases), function(i) {
  create_dfg_distmap(phases[i], phase_titles[i])
})

rm(create_fgd_distmap)
gc()


```

```{r dfgco2_dfm, message=FALSE}

# calculating mean max mixing depth across all 3 phases for each grid cell
surface_data[, ":=" (
  dFG_mean = mean(dFG, na.rm = TRUE)
), by = .(lat, lon, month)]

# can rerun code for months of choice, subset coordinates in place
create_dfg_meanmap <- function(phase_name, title_text) {
  phase_dt <- surface_data[surface_data$phase == phase_name & surface_data$month == 1]
  plot <- ggplot(phase_dt, aes(x = lon, y = lat, fill = dFG - dFG_mean)) +
    geom_polygon(data = map_data("world"), aes(x = long, y = lat, group = group), 
                 fill = "lightgray", color = "white") +
    geom_raster() +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,
                        limits = c(-0.00005, 0.00005)) + 
                        # breaks = c(-0.000005, -0.0000025, 0, 0.0000025, 0.000005)) +
    theme_light() +
    coord_fixed(xlim = c(-135, -100),
              ylim = c(10, 45)) +
    scale_x_continuous(breaks = seq(-135, -100, by = 10)) +
    scale_y_continuous(breaks = seq(10, 45, by = 10)) +
    labs(x = "Longitude",
         y = "Latitude",
         fill = "Difference from Mean dCO2 Flux 
         (mmol/m^2/s)",
         title = paste0(title_text, " Phase, Difference from Mean delta Air-Sea CO2 Flux (Month 1)")) +
    theme(panel.border = element_blank())
  
  print(plot)

  # # save plot
  # ggsave(paste0(path_plots, phase_name, "_dfgmapdfm_month1.png"), plot = plot,
  #        width = 8, height = 6, dpi = 300)
}

lapply(seq_along(phases), function(i) {
  create_dfg_meanmap(phases[i], phase_titles[i])
})

rm(create_dfg_meanmap)
gc()

```

```{r co2_uptake, message=FALSE}

# merging area data to dFG and calculating moles per month from the dFG rate
uptake_data <- surface_data[, .SD, .SDcols = c("lat", "lon", "time", "dFG", 
                                               "month", "phase")] %>% 
  merge(lanina_dTA_mol, by = c("lon", "lat"), all.x = TRUE) 

uptake_data[, secs := as.numeric(days_in_month(ymd(paste0(time, "-01"))))
            * 60 * 60 * 24] # pulls days in the month and converts to seconds
uptake_data[, CO2_uptake := dFG * area * secs / 1000] # mmol/m2/s -> moles

# can rerun code for months of choice, using zoomed region
create_uptake_distmap <- function(phase_name, title_text) {
  phase_dt <- uptake_data[uptake_data$phase == phase_name & uptake_data$month == 24]
  plot <- ggplot(phase_dt, aes(x = lon, y = lat, fill = CO2_uptake)) +
    geom_polygon(data = map_data("world"), aes(x = long, y = lat, group = group), 
                 fill = "lightgray", color = "white") +
    geom_raster() +
    # scale_fill_viridis_c(limit = c(0, 91000000)) +
    scale_fill_gradient2(low = "green", mid = "white", high = "purple", midpoint = 0,
                         limit = c(0, 1500000)) +
    theme_light() +
    coord_fixed(xlim = c(-135, -100),
              ylim = c(10, 45)) +
    scale_x_continuous(breaks = seq(-135, -100, by = 10)) +
    scale_y_continuous(breaks = seq(10, 45, by = 10)) +
    labs(x = "Longitude",
         y = "Latitude",
         fill = "CO2 Uptake (moles)",
         title = paste0(title_text, " Phase, CO2 Uptake (Month 24)")) +
    theme(panel.border = element_blank())
  
  print(plot)

  # # save plot
  # ggsave(paste0(path_plots, phase_name, "_co2uptakemap_month24.png"), plot = plot,
  #        width = 8, height = 6, dpi = 300)
}

lapply(seq_along(phases), function(i) {
  create_uptake_distmap(phases[i], phase_titles[i])
})

# integrate by time and compare it to the native grid dDIC to check scope
# # need to add these up sequentially
uptake_data_int <- uptake_data[, .(CO2_sum_ind = sum(CO2_uptake, na.rm = TRUE)),
                                   by = .(phase, month)] %>% 
  .[, CO2_sum_cum := cumsum(CO2_sum_ind), by = phase]

uptake_data_int <- merge(uptake_data_int, full_cdreff_og[, .(phase, month, dDIC_sum)], 
                         by = c("phase", "month"), all.x = TRUE) %>% 
  .[, fraction := CO2_sum_cum / dDIC_sum]

# saving uptake individual and sum value tables
write_feather(uptake_data, paste0(path_outputs,
                                      "fguptake_data.feather"))
write_feather(uptake_data_int, paste0(path_outputs,
                                      "fguptake_data_intsum.feather"))

# plot time series of accumulating uptaken CO2
ggplot(uptake_data_int, aes(x = month, y = CO2_sum_cum/1e10, color = phase)) +
  scale_x_continuous(breaks = seq(0, 24, by = 3), limits = c(0,24),
                     expand = c(0,0)) +
  scale_y_continuous(limits = c(0, 18), expand = c(0,0)) +
  geom_vline(xintercept = seq(0, 24, by = 12), color = "grey75", linewidth = 0.4) +
  geom_hline(yintercept = 16.7, linetype = "dashed", color = "black", linewidth = 0.6) +
  geom_line(linewidth = 1) +
  labs(title = "Cumulative Uptaken CO2 from Surface Flux Values", 
       x = "Months Since OAE Began", 
       y = "Moles of CO2 (10^10)") +
  theme_bw()

# ggsave(paste0(path_plots, "total_fg_co2uptake.png"), plot = last_plot(),
#               width = 10, height = 6, dpi = 300)



rm(list = ls())
gc()

```

