---
title: "Hovmoeller dTA, dDIC, and CDR Efficiency Plots"
author: "Victoria Froh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Introduction

1.  Plotting dTA, dDIC, and CDR efficiency across depth and time of the OAE addition site.
2.  Also plotting their differences from a mean of all 3 phases to compare.

```{r setup, message=FALSE}

#loading packages
library(tidyverse)
library(data.table)
library(arrow)
library(scales)

# Path to intermediate computation outputs
path_outputs <- "/net/sea/work/vifroh/oae_ccs_roms_data/regrid_2/"

# Path to save practice plots when working on them
path_plots <- "/net/sea/work/vifroh/test_plots/"

# loading in previous saved depth integrated data for hovmo plots
lanina_depthint <- read_feather(
  paste0(path_outputs, "lanina_depthintRG2.feather"))
neutral_depthint <- read_feather(
  paste0(path_outputs, "neutral_depthintRG2.feather"))
elnino_depthint <- read_feather(
  paste0(path_outputs, "elnino_depthintRG2.feather"))

# plotting
phase_titles <- c("La Niña", "Neutral", "El Niño")
phases <- c("lanina", "neutral", "elnino")

# loading in hbls weighted mean data
hbls_wmeans <- read_feather(paste0(path_outputs, "hbls_wmeansRG2.feather"))

```

# Regular Hovmoeller Plots

## Change in total (added) alkalinity

```{r  dTA_hovmo_plots, message=FALSE}

# formatting tables, subsetting
phase_data <- list(lanina_depthint, neutral_depthint, elnino_depthint)

phase_data <- phase_data %>%
  lapply(function(table) {
    setDT(table)
    table[, depth := as.numeric(depth)]
    table[, dTA_sum := as.numeric(as.character(dTA_sum))]
    table[, CDR_eff := as.numeric(as.character(CDR_eff))]
    table[, month := .GRP, by = time] # gives index in order to each unique time
    table <- table[month <= 24] # subsets to first two years
    return(table)
  })

# return tables to each item 
lanina_depthint_sub <- phase_data[[1]]
neutral_depthint_sub <- phase_data[[2]]
elnino_depthint_sub <- phase_data[[3]]

# # removing negative data points and transform
# lanina_depthint_sub <- lanina_depthint_sub[dTA_sum > 0]
# lanina_depthint_sub[, dTA_log := log10(dTA_sum)]

# setting boundaries of bins
depth_ranges <- unique(lanina_depthint_sub[, .(depth)])
setorder(depth_ranges, depth)

depth_ranges[, ":=" (
    depth_upper = (shift(depth, type = "lag", fill = 0) + depth) / 2, # set shallower boundary
    depth_lower = (shift(depth, type = "lead", fill = max(depth)) + depth) / 2 
  )] # deeper bound

# merge depth ranges back to tables and divide into surface and subsurface
phase_data <- list(lanina_depthint_sub, neutral_depthint_sub, elnino_depthint_sub)

phase_data <- phase_data %>%
  lapply(function(table) {
    table <- merge(table, depth_ranges, by = "depth", all.x = TRUE)
    table[, category := fifelse(depth < 100, "Surface", "Subsurface")]
    table$category <- factor(table$category, levels = c("Surface", "Subsurface"))
    return(table)
  })

# return tables to each item
lanina_depthint_sub <- phase_data[[1]]
neutral_depthint_sub <- phase_data[[2]]
elnino_depthint_sub <- phase_data[[3]]

# plotting
create_hovmo_dTA <- function(data, title_text, phase) {
  plot <- ggplot(data, aes(x = month, y = depth, fill = dTA_sum)) +
    geom_rect(aes(ymin = depth_upper, ymax = depth_lower, 
                  xmin = month - 1, xmax = month)) +
    scale_x_continuous(breaks = seq(0, 24, by = 3), limits = c(0,24),
                       expand = c(0,0)) + 
    scale_y_reverse(expand = c(0, 0)) +
    scale_fill_viridis_c(guide = guide_colorbar(
      barwidth = unit(0.7, "cm"), barheight = unit(6, "cm")),
      limits = c(-926000, 3443000000)) +
      # limits = c(0, 3443000000)) +
    facet_wrap(~category, scales = "free_y", ncol = 1) + # splits into two panels
    labs(x = "Months Since OAE Start",
         y = "Depth (m)",
         fill = "Added Alkalinity\n(mol/m)",
         title = paste0(title_text, " Phase Added Alkalinity")) + 
    theme_bw() + 
    theme(strip.text = element_blank(),
          axis.ticks.length = unit(0.15, "inches"),
          panel.border = element_blank(),  # Remove box around facets
          panel.grid = element_blank(),
          panel.spacing = unit(0.5, "lines"),
          strip.background = element_blank())
  
  # print(plot)
  
  # # save plot
  # ggsave(paste0(path_plots, phase, "_hovmo_dTA.png"), plot = plot,
  #        width = 8, height = 6, dpi = 300)
}

lapply(seq_along(phase_data), function(i) {
  create_hovmo_dTA(phase_data[[i]], phase_titles[i], phases[i])
})

rm(create_hovmo_dTA, lanina_depthint, elnino_depthint, neutral_depthint)
gc()

```

## Change in DIC

```{r  dDIC_hovmo_plots, message=FALSE}

# building off of data ran in chunk above

# plotting dDIC
create_hovmo_dDIC <- function(data, title_text, phase) {
  plot <- ggplot(data, aes(x = month, y = depth, fill = dDIC_sum)) +
    geom_rect(aes(ymin = depth_upper, ymax = depth_lower, 
                  xmin = month - 1, xmax = month)) +
    scale_y_reverse(expand = c(0, 0)) +
    scale_x_continuous(breaks = seq(0, 24, by = 3), limits = c(0,24),
                       expand = c(0,0)) +
    scale_fill_viridis_c(guide = guide_colorbar(
      barwidth = unit(0.7, "cm"), barheight = unit(6, "cm")),
      limits = c(-480, 2008000000)) +
      # limits = c(0, 2008000000)) +
    facet_wrap(~category, scales = "free_y", ncol = 1) + # splits into two panels
    labs(x = "Months Since OAE Start",
         y = "Depth (m)",
         fill = "Change in DIC\n(mol/m)",
         title = paste0(title_text, " Phase Change in DIC")) + 
    theme_bw() + 
    theme(strip.text = element_blank(),
          axis.ticks.length = unit(0.15, "inches"),
          panel.border = element_blank(),  # Remove box around facets
          panel.grid = element_blank(),
          panel.spacing = unit(0.5, "lines"),
          strip.background = element_blank())
  
  # print(plot)
  
  # # save plot
  # ggsave(paste0(path_plots, phase, "_hovmo_dDIC.png"), plot = plot,
  #        width = 8, height = 6, dpi = 300)
}

lapply(seq_along(phase_data), function(i) {
  create_hovmo_dDIC(phase_data[[i]], phase_titles[i], phases[i])
})

rm(create_hovmo_dDIC)
gc()

```

## Integrated CDR Efficiency (full total dDIC per layer/full total dTA per layer)

```{r  CDReff_hovmo_plots, message=FALSE}

# building off of data ran in chunk above

# plotting CDReff
create_hovmo_CDReff <- function(data, title_text, phase) {
  plot <- ggplot(data, aes(x = month, y = depth, fill = CDR_eff)) +
    geom_rect(aes(ymin = depth_upper, ymax = depth_lower, 
                  xmin = month - 1, xmax = month)) +
    scale_y_reverse(expand = c(0, 0)) +
    scale_x_continuous(breaks = seq(0, 24, by = 3), limits = c(0,24),
                       expand = c(0,0)) +
    scale_fill_viridis_c(guide = guide_colorbar(
      barwidth = unit(0.7, "cm"), barheight = unit(6, "cm")),
      limits = c(0, 1)) + # cutting off weird negative or >1 CDR effs
    facet_wrap(~category, scales = "free_y", ncol = 1) + # splits into two panels
    labs(x = "Months Since OAE Start",
         y = "Depth (m)",
         fill = "CDR Efficiency\n(Fraction)",
         title = paste0(title_text, " Phase CDR Efficiency")) + 
    theme_bw() + 
    theme(strip.text = element_blank(),
          axis.ticks.length = unit(0.15, "inches"),
          panel.border = element_blank(),  # Remove box around facets
          panel.grid = element_blank(),
          panel.spacing = unit(0.5, "lines"),
          strip.background = element_blank())
  
  # print(plot)
  
  # # save plot
  # ggsave(paste0(path_plots, phase, "_hovmo_CDReff.png"), plot = plot,
  #        width = 8, height = 6, dpi = 300)
}

lapply(seq_along(phase_data), function(i) {
  create_hovmo_CDReff(phase_data[[i]], phase_titles[i], phases[i])
})

rm(create_hovmo_CDReff)
gc()

```

# Hovmoeller plots of the differences from the means of all 3 phases

## Difference in added alkalinity from the phases' mean

```{r  delta_hovmo_dTA, message=FALSE}

# building off of data ran in chunk above

# combining all three tables in to one
phase_list <- Map(function(table, phase) {
    table[, phase := phase]
    table[, .(month, depth, depth_upper, depth_lower, category, dTA_sum, dDIC_sum, 
              CDR_eff, phase)]
  }, phase_data, phases)

phase_fulldata <- rbindlist(phase_list)

# calculating mean dTA, dDIC, and CDR-eff across all 3 phases
phase_fulldata[, ":=" (
  dTA_mean = mean(dTA_sum, na.rm = TRUE),
  dDIC_mean = mean(dDIC_sum, na.rm = TRUE),
  CDR_effmean = mean(CDR_eff[CDR_eff >= 0 & CDR_eff <= 1] , na.rm = TRUE)
), by = .(month, depth)]

#saving data table with all 3 phases and the means
# write_feather(phase_fulldata, paste0(path_outputs, "full_depthint_means.feather"))

hbls_wmeans <- rbind(hbls_wmeans, hbls_wmeans[(.N-2):.N][, month := 25])
# plotting delta dTA from mean
create_hovmo_ddTA <- function(phase_name, title_text) {
  phase_dt <- phase_fulldata[phase_fulldata$phase == phase_name,]
  hbls_dt <- hbls_wmeans[hbls_wmeans$phase == phase_name]
  plot <- ggplot() +
    geom_rect(data = phase_dt, aes(xmin = month - 1, xmax = month,
                  ymin = depth_upper, ymax = depth_lower, fill = dTA_sum - dTA_mean)) +
    scale_y_reverse(expand = c(0, 0), limits = c(190,0), 
                    sec.axis = dup_axis(labels = NULL)) +
    scale_x_continuous(breaks = seq(3, 24, by = 3), limits = c(0,24),
                       expand = c(0,0), sec.axis = dup_axis(labels = NULL)) +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,
                         guide = guide_colorbar(barwidth = 20, barheight = 1, 
                                                title.position = "bottom", title.hjust = 0.5),
      limits = c(-410000000, 430000000)) +
    geom_path(data = hbls_dt, aes(x = month - 1, y = hbls_wmean),
            color = "black", size = 1) +
    # facet_wrap(~category, scales = "free_y", ncol = 1) + # splits into two panels
    labs(x = "Months Since OAE Start",
         y = "Depth [m]",
         fill = "\u0394TA' [mol/m]",
         # title = paste0(title_text, " Phase, Added Alkalinity Difference from Mean")
         ) +
    theme_bw() +
    theme(strip.text = element_blank(),
        axis.text = element_text(size = 17, color = "black"),
        axis.title = element_text(size = 18),
        axis.ticks = element_line(size = 0.5),
        axis.ticks.y.right = element_line(size = 0.5),  # Add ticks to right side
        axis.ticks.x.top = element_line(size = 0.5),
        axis.ticks.length = unit(0.2, "cm"),
        axis.title.x.top = element_blank(),  # Remove x-axis title on top
        axis.title.y.right = element_blank(),
        # panel.border = element_blank(),  # Remove box around facets
        panel.grid = element_blank(),
        panel.spacing = unit(0.5, "lines"),
        strip.background = element_blank(),
        legend.position = "bottom",
        legend.direction = "horizontal",
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 16)
        ) +
  coord_cartesian(clip = 'off', expand = FALSE)
  
  # print(plot)
  
  # # save plot
  # ggsave(paste0(path_plots, phase_name, "_hovmo_ddTA.png"), plot = plot,
  #        width = 12, height = 8, dpi = 300)
}

lapply(seq_along(phases), function(i) {
  create_hovmo_ddTA(phases[i], phase_titles[i])
})

rm(phase_list, create_hovmo_ddTA)
gc()

```

## Difference in dDIC from the phases' mean

```{r  delta_hovmo_dDIC, message=FALSE}

# building off of data ran in chunk above

# plotting delta dDIC from mean
create_hovmo_ddDIC <- function(phase_name, title_text) {
  phase_dt <- phase_fulldata[phase_fulldata$phase == phase_name,]
  hbls_dt <- hbls_wmeans[hbls_wmeans$phase == phase_name]
  plot <- ggplot() +
    geom_rect(data = phase_dt, aes(xmin = month - 1, xmax = month,
                  ymin = depth_upper, ymax = depth_lower, fill = dDIC_sum - dDIC_mean)) +
    scale_x_continuous(breaks = seq(0, 24, by = 3), limits = c(0,24),
                       expand = c(0,0)) +
    scale_y_reverse(expand = c(0, 0), limits = c(190,0), sec.axis = dup_axis(labels = NULL)) +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,
                         guide = guide_colorbar(
      barwidth = unit(0.7, "cm"), barheight = unit(6, "cm")),
      limits = c(-200000000, 200000000)) +
    facet_wrap(~category, scales = "free_y", ncol = 1) + # splits into two panels
    labs(x = "Months Since OAE Start",
         y = "Depth (m)",
         fill = "dDIC Difference\n(mol/m)",
         title = paste0(title_text, " Phase, dDIC Difference from Mean")) +
    theme_bw() +
    theme(strip.text = element_blank(),
          axis.ticks.length = unit(0.15, "inches"),
          panel.border = element_blank(),  # Remove box around facets
          panel.grid = element_blank(),
          panel.spacing = unit(0.5, "lines"),
          strip.background = element_blank())
 # print(plot)
  
  # # save plot
  # ggsave(paste0(path_plots, phase_name, "_hovmo_ddDIC.png"), plot = plot,
  #        width = 8, height = 6, dpi = 300)
}

lapply(seq_along(phases), function(i) {
  create_hovmo_ddDIC(phases[i], phase_titles[i])
})

rm(create_hovmo_ddDIC)
gc()

```

## Difference in CDR efficiency from the phases' mean.

Wonky efficiencies outside the range from 0 to 1 were filtered out (primarily below 150m)

```{r  delta_hovmo_CDReff, message=FALSE}

# building off of data ran in chunk above; CDR efficiencies outside 0-1 were
# excluded from the mean calculation since they are casued by errors

# plotting delta dDIC from mean
create_hovmo_dCDReff <- function(phase_name, title_text) {
  phase_dt <- phase_fulldata[phase_fulldata$phase == phase_name,]
  hbls_dt <- hbls_wmeans[hbls_wmeans$phase == phase_name]
  # filter out negative dTA/dDIC/wrong CDReff as these will make weird/incorrect efficiency
  phase_dt <- phase_dt[dTA_sum < 0, dTA_sum := NA] 
  phase_dt <- phase_dt[dDIC_sum < 0, dDIC_sum := NA]
  phase_dt <- phase_dt[CDR_eff < 0 | CDR_eff > 1, CDR_eff := NA]
  phase_dt <- phase_dt[depth < 200]
  
  plot <- ggplot() +
    geom_rect(data = phase_dt, aes(xmin = month - 1, xmax = month,
                  ymin = depth_upper, ymax = depth_lower, fill = CDR_eff - CDR_effmean)) +
    scale_x_continuous(breaks = seq(0, 24, by = 3), limits = c(0,24),
                       expand = c(0,0), sec.axis = dup_axis(labels = NULL)) +
    scale_y_reverse(expand = c(0, 0), limits = c(190,0), 
                    sec.axis = dup_axis(labels = NULL)) +
    scale_fill_gradient2(na.value = "lightgray", low = "blue", mid = "white", 
                         high = "red", midpoint = 0,
                         guide = guide_colorbar(
      barwidth = 20, barheight = 1, title.position = "bottom", title.hjust = 0.5),
      limits = c(-0.152, 0.152)) +
    geom_path(data = hbls_dt, aes(x = month - 1, y = hbls_wmean),
            color = "black", size = 1) +
    # facet_wrap(~category, scales = "free_y", ncol = 1) + # splits into two panels
    labs(x = "Months Since OAE Start",
         y = "Depth [m]",
         fill = "\u0394\u03B7",
         # title = paste0(title_text, " Phase, CDR Efficiency Difference from Mean")
         ) +
    theme_bw() +
    theme(strip.text = element_blank(),
        axis.text = element_text(size = 17, color = "black"),
        axis.title = element_text(size = 18),
        axis.ticks = element_line(size = 0.5),
        axis.ticks.y.right = element_line(size = 0.5),  # Add ticks to right side
        axis.ticks.x.top = element_line(size = 0.5),
        axis.ticks.length = unit(0.2, "cm"),
        axis.title.x.top = element_blank(),  # Remove x-axis title on top
        axis.title.y.right = element_blank(),
        # panel.border = element_blank(),  # Remove box around facets
        panel.grid = element_blank(),
        panel.spacing = unit(0.5, "lines"),
        strip.background = element_blank(),
        legend.position = "bottom",
        legend.direction = "horizontal",
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 16)
        ) +
  coord_cartesian(clip = 'off', expand = FALSE)
  
  # # save plot
  # ggsave(paste0(path_plots, phase_name, "_hovmo_dCDReff_200.png"), plot = plot,
  #        width = 12, height = 8, dpi = 300)
}

lapply(seq_along(phases), function(i) {
  create_hovmo_dCDReff(phases[i], phase_titles[i])
})

rm(list = ls())
gc()

```
